Music Versus No Music Effectiveness On Cognitive Response Time and Typing Efficiency

### 2024-03-31 paper/paper.tex

\author{Shea Spalding}
\email{Shea.Spalding@colostate.edu} %update email 
\affiliation{%
  \institution{Colorado State University}
  \city{Fort Collins}
  \state{Colorado}
  \country{USA}
  \postcode{80523-1873} %chktex 8
}

@@

\section{Related Works}
While there is not necessarily a concrete answer for whether music or background noise is effective on cognitive function, there have been many
studies that attempt to determine the answer. The other studies attempted to search for this answer by
focusing on word processing on a computer with the background music and the user's ability to formulate sentences~\cite{ransdell2001141}, 
or determining if background music help users to keep their attention on their task at hand~\cite{kiss2021effect}. Similar to~\cite{kiss2021effect},~\cite{mathew2022inherently} also examined how motivation and music were used hand-in-hand on ``a time-consuming task and task performance''. A different study done in 2010 again found similar results of music both being a hinderance and benefit~\cite{AudioDistractionsAshley}. According to~\cite{Gonzalez_Aiello_2019}, the degree of cognitive function may be based on an individual's personality.

Although~\cite{ransdell2001141} suggests that the music is a detriment, those with ``musical training and high working memory span wrote better essays with longer sentences'. This suggests that while the music slowed down the participants, it did not necessarily lower their cognitive ability in typing on their computer while focusing on forming text for the written essay for this particular study. For~\cite{AudioDistractionsAshley}, the music caused enough of a difference in the time that was taken for writing a paragraph but ``there were no signigicant differences in typing speed between the audio distractions''. This again may suggest that the participants cognitive ability was not hindered when typing and may transfer to directly after should the participants taken a test measuring their cognitive function. The slow timing was again seen in~\cite{Gonzalez_Aiello_2019} in which ``music generally impaired performance on a complex task but improved performance on a simple task''.

The study done through~\cite{kiss2021effect} saw opposite results as to~\cite{Gonzalez_Aiello_2019}. The music or background noise ``enchanced'' the reaction time and performance of the tasks that we given to be completed. The results from~\cite{mathew2022inherently} were split. The music was a detriment for logic puzzles in the user's accuracy but aided in the efficiency of a writing task.

\section{Methodology}
  After obtaining consent from all participants, they proceeded to take the typing test. The test presented lyrics from popular songs within a graphical user interface (GUI). 
  The interface automatically calculated typing speed and accuracy in the program's backend, allowing participants to focus soley on completing the test.
  Each participant completed the typing test under three conditions: music with lyrics, music without lyrics, and no music (silent condition).
  Multiple levels of music intensity were utilized within each condition to assess varying degrees of distraction.
  The GUI of the typing test was built with simplicity in mind. The starting window was pulled up so as not to confuse participants with the code as seen in Figure 1. The participants saw Figure 2 while the test is running, while Figure 3 is the signifier that the typing portion is finished.

@@

  They were given the instructions in the GUI with the text area where their text goes clearly visible and the only button in the window to start the test and timer. After they hit the ``Get the Typing Test'' button, the lyrics of the randomly selected songs were displayed for the user to type. The text box for the users was already in focus with a ``Done'' button. After the users hit ``Done'', their accuracy and speed were presented.

  Following the typing test, participants proceeded to to the cognitive task component of the study, which were also conducted under the three aforementioned conditions. While the participants did the cognitive task of a quantified behavioral test, they wore the EyeWriter 2.0 glasses that we specially built for this study.
  % Further details on the cognitive task will be provided in subsequent sections. They will then move on the the QBTest?. % This will need to be re-written.

>>>>

\author{Shea Spalding}
\email{Shea1@colostate.edu} %update email 
\affiliation{%
  \institution{Colorado State University}
  \city{Fort Collins}
  \state{Colorado}
  \country{USA}
  \postcode{80521-1873} %chktex 8
}

@@

\section{Related Works}
While there is not necessarily a concrete answer for whether music or background noise is effective on cognitive function, there have been many
studies that attempt to determine the answer. The other studies attempted to search for this answer by
focusing on word processing on a computer with the background music and the user's ability to formulate sentences~\cite{ransdell2001141}, 
or determining if background music helps users to keep their attention on their task at hand~\cite{kiss2021effect}. Similar to~\cite{kiss2021effect},~\cite{mathew2022inherently} also examined how motivation and music were used hand-in-hand on ``a time-consuming task and task performance''. A different study done in 2010 again found similar results of music both being a hindrance and benefit~\cite{AudioDistractionsAshley}. According to~\cite{Gonzalez_Aiello_2019}, the degree of cognitive function may be based on an individual's personality.

Although~\cite{ransdell2001141} suggests that the music is a detriment, those with ``musical training and high working memory span wrote better essays with longer sentences'. This suggests that while the music slowed down the participants, it did not necessarily lower their cognitive ability in typing on their computer while focusing on forming text for the written essay for this particular study. For~\cite{AudioDistractionsAshley}, the music caused enough of a difference in the time that was taken for writing a paragraph but ``there were no significant differences in typing speed between the audio distractions''. This again may suggest that the participant's cognitive ability was not hindered when typing and may transfer directly after should the participants have taken a test measuring their cognitive function. The slow timing was again seen in~\cite{Gonzalez_Aiello_2019} in which ``music generally impaired performance on a complex task but improved performance on a simple task''.

The study done through~\cite{kiss2021effect} saw opposite results as to~\cite{Gonzalez_Aiello_2019}. The music or background noise ``enhanced'' the reaction time and performance of the tasks that are given to be completed. The results from~\cite{mathew2022inherently} were split. The music was a detriment for logic puzzles in the user's accuracy but aided in the efficiency of a writing task.

\section{Methodology}
After obtaining consent from all participants, they proceeded to take the typing test. The test presented lyrics from popular songs within a graphical user interface (GUI). 
The interface automatically calculated typing speed and accuracy in the program's backend, allowing participants to focus solely on completing the test.
Each participant completed the typing test under three conditions: music with lyrics, music without lyrics, and no music (silent condition).
Multiple levels of music intensity were utilized within each condition to assess varying degrees of distraction.
The GUI of the typing test was built with simplicity in mind. The starting window was pulled up so as not to confuse participants with the code as seen in Figure 1. The participants saw Figure 2 while the test is running, while Figure 3 is the signifier that the typing portion is finished.

@@

They were given the instructions in the GUI with the text area where their text is clearly visible and the only button in the window to start the test and timer. After they hit the ``Get the Typing Test'' button, the lyrics of the randomly selected songs were displayed for the user to type. The text box for the users was already in focus with a ``Done'' button. After the users hit ``Done'', their accuracy and speed were presented.

Following the typing test, participants proceeded to the cognitive task component of the study, which was also conducted under the three aforementioned conditions. While the participants did the cognitive task of a quantified behavioral test, their eye movement will be tracked by the EyeWriter 2.0 camera that we specially built for this study. The EyeWriter system consists of a camera mounted on the participant's laptop, which tracks eye movements and allows for precise measurement of gaze direction and fixation duration.

Participants were seated in a comfortable position facing the laptop screen. They were instructed to follow a series of calibration points displayed on the screen to calibrate the eye-tracking system. This calibration process ensured accurate measurement of eye movements during the subsequent tasks.

During the cognitive task component of the study, participants performed a quantified behavioral test (QBTest) while their eye movements were tracked by the EyeWriter 2.0 camera mounted on the laptop. The QBTest is a computerized test designed to assess various aspects of cognitive function, including attention, impulsivity, and response control.
% Further details on the cognitive task will be provided in subsequent sections. They will then move on the the QBTest?. % This will need to be re-written.
